---
title: "smoothedComparison"
author: "Alexandra DiGiacomo"
date: "10/15/2020"
output: html_document
---


```{r, message = FALSE}
# load packages

library(dplyr)
library(ggplot2)
library(raster)
library(sp)
library(rgdal)
library(spatialEco)
library(rasterVis)
library(rgl) 
```

## Load DSMs
```{r , message = FALSE}
# load smoothed data
IL5 <- raster("C:/aed41/Projects/OysterMetrics/data/Products/smoothed/1L5_best_dsm.tif")
CCA3 <- raster("C:/aed41/Projects/OysterMetrics/data/Products/smoothed/CCA3_Final_dsm.tif")
Natural <- raster("C:/aed41/Projects/OysterMetrics/data/Products/smoothed/Natural_low_test_dsm.tif")


# load nonsmoothed data
IL5_ns <- raster("C:/aed41/Projects/OysterMetrics/data/Products/non-smoothed/1L5_best_dsm_nonsmoothed.tif")
CCA3_ns <- raster("C:/aed41/Projects/OysterMetrics/data/Products/non-smoothed/CCA3_Final_dsm_nonsmoothed.tif")
Natural_ns <- raster("C:/aed41/Projects/OysterMetrics/data/Products/non-smoothed/Natural_low_test_dsm_nonsmoothed.tif")
```

## Load Orthos
```{r}
IL5_ortho <- brick("C:/aed41/Projects/OysterMetrics/data/Products/smoothed/1L5_best_transparent_mosaic_group1.tif")
Natural_ortho <- brick("C:/aed41/Projects/OysterMetrics/data/Products/smoothed/Natural_Low_test_transparent_mosaic_group1.tif")
CCA3_ortho <- brick("C:/aed41/Projects/OysterMetrics/data/Products/smoothed/CCA3_Final_transparent_mosaic_group1.tif")
```


# Generate Downsampled Rasters
```{r, cache = TRUE, include = FALSE}
# smoothed
IL52 <- aggregate(IL5, fact = 2, fun = "mean")
IL54 <- aggregate(IL5, fact = 4, fun = "mean")
IL58 <- aggregate(IL5, fact = 8, fun = "mean")
IL516 <-aggregate(IL5, fact = 16, fun = "mean")

CCA32 <- aggregate(CCA3, fact = 2, fun = "mean")
CCA34 <- aggregate(CCA3, fact = 4, fun = "mean")
CCA38 <- aggregate(CCA3, fact = 8, fun = "mean")
CCA316 <-aggregate(CCA3, fact = 16, fun = "mean")

Natural2 <- aggregate(Natural, fact = 2, fun = "mean")
Natural4 <- aggregate(Natural, fact = 4, fun = "mean")
Natural8 <- aggregate(Natural, fact = 8, fun = "mean")
Natural16 <-aggregate(Natural, fact = 16, fun = "mean")

# nonsmoothed
IL52_ns <- aggregate(IL5_ns, fact = 2, fun = "mean")
IL54_ns <- aggregate(IL5_ns, fact = 4, fun = "mean")
IL58_ns <- aggregate(IL5_ns, fact = 8, fun = "mean")
IL516_ns <-aggregate(IL5_ns, fact = 16, fun = "mean")

CCA32_ns <- aggregate(CCA3_ns, fact = 2, fun = "mean")
CCA34_ns <- aggregate(CCA3_ns, fact = 4, fun = "mean")
CCA38_ns <- aggregate(CCA3_ns, fact = 8, fun = "mean")
CCA316_ns <-aggregate(CCA3_ns, fact = 16, fun = "mean")

Natural2_ns <- aggregate(Natural_ns, fact = 2, fun = "mean")
Natural4_ns <- aggregate(Natural_ns, fact = 4, fun = "mean")
Natural8_ns <- aggregate(Natural_ns, fact = 8, fun = "mean")
Natural16_ns <-aggregate(Natural_ns, fact = 16, fun = "mean")
```

# Calculate Slope and Aspect
```{r}
# generate iteratable lists
list = c(IL5, IL52, IL54, IL58, IL516, IL5_ns, IL52_ns, IL54_ns, IL58_ns, IL516_ns,
         CCA3, CCA32, CCA34, CCA38, CCA316, CCA3_ns, CCA32_ns, CCA34_ns, CCA38_ns, CCA316_ns,
         Natural, Natural2, Natural4, Natural8, Natural16, Natural_ns, Natural2_ns, Natural4_ns,
         Natural8_ns, Natural16_ns)
names = c('IL5', 'IL52', 'IL54', 'IL58', "IL516", "IL5_ns", "IL52_ns", "IL54_ns", 
          "IL58_ns", "IL516_ns",
         "CCA3", "CCA32", "CCA34", "CCA38", 'CCA316', "CCA3_ns", "CCA32_ns", 
         "CCA34_ns", "CCA38_ns", "CCA316_ns",
         "Natural", "Natural2", "Natural4", "Natural8", "Natural16", 
         "Natural_ns", "Natural2_ns", "Natural4_ns", "Natural8_ns", "Natural16_ns")

newnames = c() # initialize empty list for new slopeAs objects

# loop
i = 1
for (i in 1:length(list)) {
  nam = paste(names[i], "slopAs", sep = "")

  rast = list[[i]] # grab name of raster
  
  slopeAs <- terrain(rast, opt = c("slope", "aspect"), unit = "degrees") # calc slope/aspect
  assign(nam, slopeAs)
  
  newnames <- append(newnames, nam)
  i = i + 1 # move pointer
}

```

# Plot Slope and Aspect
```{r, fig.height = 3, fig.width = 5}
i = 1
for (i in 1:length(newnames)) {
  jpeg(file = paste("C:/aed41/Projects/OysterMetrics/results/slopeAspectFigures/",
                       paste(newnames[[i]], ".png", sep = ""), sep = ""),
       width = 600, height = 400)
  plot(eval(parse(text = newnames[[i]])), main = newnames[[i]])
  dev.off()
}
```

## Create Raster Stacks of Slope/Aspect Across Resolutions/Smoothing
```{r}
rasStackIL5 = stack(IL5slopAs, resample(IL52slopAs, IL5slopAs), 
                 resample(IL54slopAs, IL5slopAs),
                 resample(IL58slopAs, IL5slopAs),
                 resample(IL516slopAs, IL5slopAs),
                 resample(IL5_nsslopAs, IL5slopAs),
                 resample(IL52_nsslopAs, IL5slopAs),
                 resample(IL54_nsslopAs, IL5slopAs),
                 resample(IL58_nsslopAs, IL5slopAs),
                 resample(IL516_nsslopAs, IL5slopAs))

rasStackNatural = stack(NaturalslopAs, resample(Natural2slopAs, NaturalslopAs), 
                 resample(Natural4slopAs, NaturalslopAs),
                 resample(Natural8slopAs, NaturalslopAs),
                 resample(Natural16slopAs, NaturalslopAs),
                 resample(Natural_nsslopAs, NaturalslopAs),
                 resample(Natural2_nsslopAs, NaturalslopAs),
                 resample(Natural4_nsslopAs, NaturalslopAs),
                 resample(Natural8_nsslopAs, NaturalslopAs),
                 resample(Natural16_nsslopAs, NaturalslopAs))

rasStackCCA3 = stack(CCA3slopAs, resample(CCA32slopAs, CCA3slopAs), 
                 resample(CCA34slopAs, CCA3slopAs),
                 resample(CCA38slopAs, CCA3slopAs),
                 resample(CCA316slopAs, CCA3slopAs),
                 resample(CCA3_nsslopAs, CCA3slopAs),
                 resample(CCA32_nsslopAs, CCA3slopAs),
                 resample(CCA34_nsslopAs, CCA3slopAs),
                 resample(CCA38_nsslopAs, CCA3slopAs),
                 resample(CCA316_nsslopAs, CCA3slopAs))
```


# Vertical Height Profile Check
Traverse a horizontal transect across the raster (a row of the raster matrix) and visualize changes in the vertical profile across resolutions and smoothed/unsmoothed techniques

```{r}
# extract three horizontal rows for profile

# IL5
nrowIL5 <- dim(rasStackIL5)[1] # num rows in this raster
trowIL5 <- round(nrowIL5/4) # cut the raster into quarters horizontally, take middle three transects

transect1_IL5 <- as.data.frame(getValues(rasStackIL5, trowIL5, 1)) %>%
  mutate(id = row_number())
transect2_IL5 <- as.data.frame(getValues(rasStackIL5, trowIL5*2, 1)) %>%
  mutate(id = row_number())
transect3_IL5 <- as.data.frame(getValues(rasStackIL5, trowIL5*3, 1)) %>%
  mutate(id = row_number())

# Natural
nrowNatural <- dim(rasStackNatural)[1] # num rows in this raster
trowNatural <- round(nrowNatural/4) # cut the raster into quarters horizontally, take middle three transects

transect1_Natural <- as.data.frame(getValues(rasStackNatural, trowNatural, 1)) %>%
  mutate(id = row_number())
transect2_Natural <- as.data.frame(getValues(rasStackNatural, trowNatural*2, 1)) %>%
  mutate(id = row_number())
transect3_Natural <- as.data.frame(getValues(rasStackNatural, trowNatural*3, 1)) %>%
  mutate(id = row_number())

# CCA3
nrowCCA3 <- dim(rasStackCCA3)[1] # num rows in this raster
trowCCA3 <- round(nrowCCA3/4) # cut the raster into quarters horizontally, take middle three transects

transect1_CCA3 <- as.data.frame(getValues(rasStackCCA3, trowCCA3, 1)) %>%
  mutate(id = row_number())
transect2_CCA3 <- as.data.frame(getValues(rasStackCCA3, trowCCA3*2, 1)) %>%
  mutate(id = row_number())
transect3_CCA3 <- as.data.frame(getValues(rasStackCCA3, trowCCA3*3, 1)) %>%
  mutate(id = row_number())
```

Reformat Data for Plotting
```{r plotverticalProfile}
transects <- do.call(rbind, list(t1_IL5 = transect1_IL5, t2_IL5 = transect2_IL5, 
                    t3_IL5 = transect3_IL5, t1_Nat = transect1_Natural, 
                    t2_Nat = transect2_Natural, t3_Nat = transect3_Natural, 
                    t1_CCA = transect1_CCA3, t2_CCA = transect2_CCA3, t3_CCA = transect3_CCA3))

transects$row = rownames(transects) # add row name

ts <- transects %>%
  dplyr::select(!(starts_with("aspect"))) %>%
  pivot_longer(cols= starts_with("slope"), names_to = "slope", values_to = "value") %>%
  dplyr::mutate(type = dplyr::case_when(
    slope == "slope.1" ~ "1xGSD smoothed",
    slope ==  "slope.2" ~ "2xGSD smoothed",
    slope == "slope.3" ~ "4xGSD smoothed",
    slope ==  "slope.4" ~ "8xGSD smoothed",
    slope == "slope.5" ~ "16xGSD smoothed",
    slope ==  "slope.6" ~ "1xGSD unsmoothed",
    slope ==   "slope.7" ~ "2xGSD unsmoothed",
    slope ==  "slope.8" ~ "4xGSD unsmoothed",
    slope ==   "slope.9" ~ "8xGSD unsmoothed",
    slope ==   "slope.10" ~ "16xGSD unsmoothed",
  )) %>%
  separate(row, into = c("transect", "site"), sep = "_")%>%
  separate(site, into = c("site", "scratch"), sep = "([.])") %>%
  dplyr::select(-scratch) %>%
  separate(type, into = c("GSD", "smoothing"), sep = " ") %>%
  rename(slope_degrees = value) %>%
  mutate(position_cm = id*.00586*100)

ts
```

Plot: 1L5
```{r}
ts %>%
  filter(site == "IL5" & transect == "t1", ncol = 1) %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing) +
  ggtitle("1L5: Transect 1")

ts %>%
  filter(site == "IL5" & transect == "t2") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("1L5: Transect 2")

ts %>%
  filter(site == "IL5" & transect == "t3") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("1L5: Transect 3")

library(RStoolbox)
ggRGB(IL5_ortho, r = 1, g = 2, b = 3) + 
  geom_hline(yintercept = extent(IL5_ortho)[4]-(trowIL5*.0056), size = 3, color = "red") +
  geom_hline(yintercept = extent(IL5_ortho)[4]-(trowIL5*.0056)*2, size = 3, color = "red") +
  geom_hline(yintercept = extent(IL5_ortho)[4]-(trowIL5*.0056)*3, size = 3, color = "red") + 
  ggtitle("Transect Positions: 1L5")
```
Plot: Natural
```{r}
ts %>%
  filter(site == "Nat" & transect == "t1") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("Natural: Transect 1")

ts %>%
  filter(site == "Nat" & transect == "t2") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("Natural: Transect 2")

ts %>%
  filter(site == "Nat" & transect == "t3") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("Natural: Transect 3")

library(RStoolbox)
ggRGB(Natural_ortho, r = 1, g = 2, b = 3) + 
  geom_hline(yintercept = extent(Natural_ortho)[4]-(trowNatural*.0056), size = 3, color = "red") +
  geom_hline(yintercept = extent(Natural_ortho)[4]-(trowNatural*.0056)*2, size = 3, color = "red") +
  geom_hline(yintercept = extent(Natural_ortho)[4]-(trowNatural*.0056)*3, size = 3, color = "red") + 
  ggtitle("Transect Positions: Natural")
```

Plot: CCA3
```{r}
ts %>%
  filter(site == "CCA" & transect == "t1") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("CCA3: Transect 1")

ts %>%
  filter(site == "CCA" & transect == "t2") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("CCA3: Transect 2")

ts %>%
  filter(site == "CCA" & transect == "t3") %>%
  ggplot(aes( x = position_cm, y = slope_degrees, color = GSD)) + geom_line() + 
  theme_minimal() + facet_wrap(~ smoothing, ncol = 1) +
  ggtitle("CCA3: Transect 3")

library(RStoolbox)
ggRGB(CCA3_ortho, r = 1, g = 2, b = 3) + 
  geom_hline(yintercept = extent(CCA3_ortho)[4]-(trowCCA3*.0056), size = 3, color = "red") +
  geom_hline(yintercept = extent(CCA3_ortho)[4]-(trowCCA3*.0056)*2, size = 3, color = "red") +
  geom_hline(yintercept = extent(CCA3_ortho)[4]-(trowCCA3*.0056)*3, size = 3, color = "red") + 
  ggtitle("Transect Positions: CCA3")
```



Smoothed aspect (need circular statistics for this actually)
```{r}
dev.new()
plot(x = transect1$id, y = transect1$aspect.1, type = 'l', col = 'blue')
plot(x = transect1$id, y = transect1$aspect.2, type = 'l', col = 'light blue')
plot(x = transect1$id, y = transect1$aspect.3, type = 'l', col = 'blue')
plot(x = transect1$id, y = transect1$aspect.4, type = 'l', col = 'dark blue')
plot(x = transect1$id, y = transect1$aspect.5, type = 'l', col = 'green')
```

## Export Data
1. Further work will focus only on 2XGSD and 4xGSD downsampling (based on visual inspectin of graphs above, best representation of trends without losing much detail)
2. Further work will only include unsmoothed data; though it is more noisy, we want to be in control of all smoothing - may use a low pass filter later
```{r}
#1L5 rasters
writeRaster(IL52_ns, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/IL52.tif")
writeRaster(IL54_ns, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/IL54.tif")
writeRaster(IL52_nsslopAs, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/IL52_slop.tif")
writeRaster(IL54_nsslopAs, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/IL54_slop.tif")

#natural rasters
writeRaster(Natural2_ns, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/Natural2.tif")
writeRaster(Natural4_ns, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/Natural4.tif")
writeRaster(Natural2_nsslopAs, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/Natural2_slop.tif")
writeRaster(Natural4_nsslopAs, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/Natural4_slop.tif")

#cca3 rasters
writeRaster(CCA32_ns, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/CCA32.tif")
writeRaster(CCA34_ns, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/CCA34.tif")
writeRaster(CCA32_nsslopAs, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/CCA32_slop.tif")
writeRaster(CCA34_nsslopAs, "C:/aed41/Projects/OysterMetrics/scratch/intermediateFiles/CCA34_slop.tif")
```

